Preprocessing: trimming

### 2022-09-15 analysis/trimming.Rmd
removed KOT outliers
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

```{r, include = FALSE}
# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

@@

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

@@

## Histogram
### Before removing outliers
```{r}
dt_ioi_subset_all <- dt_ioi[Articulation != "NA" & !is.na(IOI)]

gghistogram(dt_ioi_subset_all, x = "IOI", color = "Articulation", bins = 150)
```

### After removing outliers
Remove outliers, which are outside of +/- 3SD separately for whether responses on boundary (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in such boundaries.

```{r}
ioi_summary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]

@@

# evaluate if it's on boundary
dt_ioi_subset_all$Boundary <- "No"
dt_ioi_subset_all[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"

# exclude ioi > +- 3SD (separately for each Boundary)
ioi_boundary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]

dt_ioi_trimmed <- data.table()
for (boundary in unique(ioi_boundary$Boundary)){
  upper_ioi <- ioi_boundary[Boundary == boundary]$Mean+3*ioi_boundary[Boundary == boundary]$SD
  lower_ioi <- ioi_boundary[Boundary == boundary]$Mean-3*ioi_boundary[Boundary == boundary]$SD
  dt_current <- dt_ioi_subset_all[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
  dt_ioi_trimmed <- rbind(dt_ioi_trimmed, dt_current)
}

@@

Remove `r nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset_all))` % of total responses

# KOT

>>>>

```{r, include = FALSE}
# define subcomponents
# for intervals (Legato/Forte, Staccato/Piano)
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

@@

# define subcomponent change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define subcomponent change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

@@

## Histogram
### Before removing outliers
```{r}
dt_ioi_subset <- dt_ioi[Articulation != "NA" & !is.na(IOI)]

gghistogram(dt_ioi_subset, x = "IOI", color = "Articulation", bins = 150)
```

### After removing outliers
Remove outliers, which are outside of +/- 3SD separately depending on whether responses are on boundaries (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in the transition points.

```{r}
ioi_summary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]

@@

# evaluate if it's on boundary
dt_ioi_subset$Boundary <- "No"
dt_ioi_subset[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"

# exclude ioi > +- 3SD (separately for each Boundary)
ioi_boundary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]

dt_ioi_trimmed <- data.table()
for (boundary in unique(ioi_boundary$Boundary)){
  upper_ioi <- ioi_boundary[Boundary == boundary]$Mean+3*ioi_boundary[Boundary == boundary]$SD
  lower_ioi <- ioi_boundary[Boundary == boundary]$Mean-3*ioi_boundary[Boundary == boundary]$SD
  dt_current <- dt_ioi_subset[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
  dt_ioi_trimmed <- rbind(dt_ioi_trimmed, dt_current)
}

@@

Remove `r nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset))` % of total responses

```{r, include = FALSE}
fwrite(dt_ioi_trimmed, file = here("trimmed", "dt_ioi_trimmed.txt"), row.names = F)
```

# KOT
```{r}
dt_kot <- dt_onset

# Offset 1 - Onset 2
dt_kot$KOT <- NA
for (row in 1:nrow(dt_kot)){
  if (row < nrow(dt_kot)){
    dt_kot$KOT[row+1] <- dt_offset$TimeStamp[row] - dt_onset$TimeStamp[row+1] # offset(n) - onset(n+1)
  }
}

# remove the first note
dt_kot <- dt_kot[RowNr != 1]

# assign a sequence number for each tone
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Articulation/Dynamics
dt_kot$Articulation <- "NA"
dt_kot$Dynamics <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_kot[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
    dt_kot[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_kot[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
    dt_kot[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
  }
}

# assign Subcomponent Change
for (number in change_1){
  dt_kot[Interval == number]$Articulation <- "LtoS"
  dt_kot[ Interval == number]$Dynamics <- "FtoP"
}
for (number in change_2){
  dt_kot[Interval == number]$Articulation <- "StoL"
  dt_kot[Interval == number]$Dynamics <- "PtoF"
}
```

## Histogram
### Before removing outliers

```{r}
dt_kot_subset <- dt_kot[Articulation != "NA" & !is.na(KOT)]

gghistogram(dt_kot_subset, x = "KOT", color = "Articulation", bins = 150)
```

### After removing outliers
```{r}
kot_subcomponent <- dt_kot_subset[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = Articulation]
dt_kot_trimmed <- data.table()
for (subcomponent in unique(dt_kot_subset$Articulation)){
  upper <- kot_subcomponent[Articulation == subcomponent]$Mean+3*kot_subcomponent[Articulation == subcomponent]$SD
  lower <- kot_subcomponent[Articulation == subcomponent]$Mean-3*kot_subcomponent[Articulation == subcomponent]$SD
  dt_current <- dt_kot_subset[Articulation == subcomponent & KOT < upper & KOT > lower]
  dt_kot_trimmed <- rbind(dt_kot_trimmed, dt_current)
}

gghistogram(dt_kot_trimmed, x = "KOT", color = "Articulation", bins = 150)
```

```{r, include = FALSE}
fwrite(dt_kot_trimmed, file = here("trimmed", "dt_kot_trimmed.txt"), row.names = F)
```
